### 2011-12-09 trunk/papers/smectics/smectics.tex
Ready to submit
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@1495 40876441-3496-0410-a613-a2fca35b2ae2

To quantify our observed morphology more closely, we use the alignment function $X$ introduced in Eq.~\ref{alignment}.
Fig.~\ref{fig3} shows snapshots of the time evolution of system A-2D at modest shear rate.
All lattice sites with $X({\mathbf r})\le0.8$ are marked in red and they obviously coincide with lamellar endpoints (edge dislocations) or other local defects.
The upper panel gives the situation at the lowest applied shear rate $\dot{\gamma}=1.25\e{-5}$, immediately following a pre-shearing protocol in which the strain rate was ramped down stepwise from higher values (see Table \ref{tab1}). While large aligned regions are present there are other regions of high defect density.
The lower panel of Fig.~\ref{fig3} shows a later state with a total strain incremented by a further $6400\%$ lattice units (i.e. by $64$ times the system size). 
Obviously many defects have annealed away and the smectic layers are now rather well aligned with the flow direction.

@@

\begin{figure}[htp!]
\centering
\includegraphics[angle=0,width=0.45\textwidth]{S_xy_tot_t_5e-4.pdf}
\includegraphics[angle=0,width=0.45\textwidth]{P_xy_chem_t_5e-4.pdf}
\includegraphics[angle=0,width=0.465\textwidth]{P_xy_defect_density.pdf}
\caption{Top: average total shear stress $\sigma_{xy}$ versus time for system A-2D: The horizontal lines are steady state time averages. For shear rate $\dot{\gamma}=1.25\e{-5}, 2.5\e{-5}, 5\e{-5}, 1\e{-4}, 2\e{-4}$, we find $\eta_{app}/\eta\equiv\sigma_{xy}/ \eta \dot{\gamma}=2.49, 2.25, 2.18, 1.93, 1.74$; centre: average order parameter contribution to the pressure tensor $-P_{xy}$ vs time for system A-2D. The inset shows data for system A-3D and shear rates $\dot{\gamma}=2\times10^{-4}, 1\times10^{-4}$ and $5\times10^{-5}$ (left to right); bottom: average chemical contribution to the pressure tensor $-P_{xy}$ vs. defect density for system A-2D.}
\label{fig7}
\end{figure}

@@

Fig.~\ref{fig8} shows averaged first normal stress differences $\langle N_1({\bf r})\rangle \equiv N_1$ versus time for system A-2D, where we report the same stepwise decreasing flow protocol as before.
For the first several steps $N_1$ decreases monotonously with decreasing shear rate.
However, at the lowest shear rates the behaviour is not monotic: $N_1$ eventually rises again. (The resulting time averages, shown as horizontal lines, do not therefore reflect true steady states.)
This somewhat surprising feature has a simple explanation. 
The time sequence of the defect density in Fig. \ref{fig6} reveals that the rise in $N_1$ is correlated with a drop in $\rho_D$, and linked to the onset of defect annihilation (which happens only at low shear rates).
When two defects meet and cancel, the local lamellar width alters by a certain amount. This causes frustration as the system is unable to snap back locally towards its equilibrium layer spacing. Hence the normal stress difference $N_1$ $\sigma_{xx}-\sigma_{yy}$ fails to decay to zero, and remains positive at large times, as appropriate for a system of metastable and (on average) compressed smectic layers. The time dependence of the normal stress difference is also anomalously slow, whereas the shear stress reaches their steady state values more quickly. The relatively large steady state value of $N_1$ and this slow dynamics may partly be a finite size effect caused by the relatively modest number of smectic layers within the simulated volume.
The normal stress in system B-2D (not shown) behaves differently: in this case $N_1$ falls monotonically during the defect annihilation phase. 
It is not clear whether this difference should be attributed to the change in $\beta, Ca$ or $Pe$ between the two systems, or indeed a combination of those changes.

>>>>

To quantify our observed morphology more closely, we use the alignment function $X$ introduced in Eq.~\ref{alignment}.
Fig.~\ref{fig3} shows snapshots of the time evolution of system A-2D at modest shear rate.
All lattice sites with $X({\mathbf r})\le0.8$ are marked in red and they obviously coincide with lamellar endpoints (edge dislocations) or other local defects.
The upper panel gives the situation at the lowest applied shear rate $\dot{\gamma}=1.25\e{-5}$, immediately following a pre-shearing protocol in which the strain rate was ramped down stepwise from higher values (see Table \ref{tab1}). While large aligned regions are present there are other regions of high defect density.
The lower panel of Fig.~\ref{fig3} shows a later state with a total strain incremented by a further $6400\%$ lattice units (i.e. by a displacement $64$ times the system size). 
Obviously many defects have annealed away and the smectic layers are now rather well aligned with the flow direction.

@@

\begin{figure}[htp!]
\centering
\includegraphics[angle=0,width=0.45\textwidth]{S_xy_tot_t_5e-4.pdf}
\includegraphics[angle=0,width=0.45\textwidth]{P_xy_chem_t_5e-4.pdf}
\includegraphics[angle=0,width=0.465\textwidth]{P_xy_defect_density.pdf}
\caption{Top: average total shear stress $\sigma_{xy}$ versus time for system A-2D: The horizontal lines are steady state time averages. For shear rate $\dot{\gamma}=1.25\e{-5}, 2.5\e{-5}, 5\e{-5}, 1\e{-4}, 2\e{-4}$, we find $\eta_{app}/\eta\equiv\sigma_{xy}/ \eta \dot{\gamma}=2.49, 2.25, 2.18, 1.93, 1.74$; centre: average order parameter contribution to the stress tensor $-P_{xy}$ vs time for system A-2D. The inset shows data for system A-3D and shear rates $\dot{\gamma}=2\times10^{-4}, 1\times10^{-4}$ and $5\times10^{-5}$ (left to right); bottom: average chemical contribution to the pressure tensor $-P_{xy}$ vs. defect density for system A-2D.}
\label{fig7}
\end{figure}

@@

Fig.~\ref{fig8} shows averaged first normal stress differences $\langle N_1({\bf r})\rangle \equiv N_1$ versus time for system A-2D, where we report the same stepwise decreasing flow protocol as before.
For the first several steps $N_1$ decreases monotonously with decreasing shear rate.
However, at the lowest shear rates the behaviour is not monotic: $N_1$ eventually rises again. (The resulting time averages, shown as horizontal lines, do not therefore reflect true steady states.)
This somewhat surprising feature has a simple explanation. 
The time sequence of the defect density in Fig. \ref{fig6} reveals that the rise in $N_1$ is correlated with a drop in $\rho_D$, and linked to the onset of defect annihilation (which happens only at low shear rates).
When two defects meet and cancel, the local lamellar width alters by a certain amount. This causes frustration as the system is unable to snap back locally towards its equilibrium layer spacing. Hence the normal stress difference $N_1=\sigma_{xx}-\sigma_{yy}$ fails to decay to zero, and remains positive at large times, as appropriate for a system of metastable and (on average) compressed smectic layers. The time dependence of the normal stress difference is also anomalously slow, whereas the shear stress reaches their steady state values more quickly. The relatively large steady state value of $N_1$ and this slow dynamics may partly be a finite size effect caused by the relatively modest number of smectic layers within the simulated volume.
The normal stress in system B-2D (not shown) behaves differently: in this case $N_1$ falls monotonically during the defect annihilation phase. 
It is not clear whether this difference should be attributed to the change in $\beta, Ca$ or $Pe$ between the two systems, or indeed a combination of those changes.
